Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-12-09 manuscript/review_response_R1.Rmd

\hrulefill
\bigskip{} \bigskip{}

\textbf{Editors Comments for the Author(s)}

Please pay close attention to the comments from Reviewer 1 regarding the validity of the findings. Once you have responded to these comments in a point-by-point cover letter and using tracked changes in the manuscript document, I will likely send your manuscript out for re-review.

\response{We provide point-by-point responses to reviewer comments below. We have taken care to address Reviewer 1's concerns (also echoed by Reviewer 2) by including the suggested sensitivity analysis (see new section in Methods, lines xxx-xxx). Unfortunately, we could not implement the suggestion of Reviewer 1 to include a VWC$\times$Treatment$\times$Year interaction in our model because our data limits the complexity of our statistical model. Within a year and treatment, each plot has the same value of volumetric water content. Therefore, we simply don't have the variation needed to fit a VWC$\times$Treatment$\times$Year interaction, which would be necessary to incorporate the effect of time on the ecosystem functional response.}

\response{The sensitivity analysis allows us to examine how the magnitude of the treatment effect changes over time. We found that the response of ANPP to drought grew over time, while the response of ANPP to irrigation is static (new panel D in Figure 3). We think this analysis improves the connection of our results to the Hierarchical Response Framework.}

@@

2. I didn't see mention of the priors on $\beta$, but I might've missed it. See Lemoine et al. (2016) - Ecology for a discussion of this issue. (very minor issue, I doubt the priors make too much difference here).

\response{Sorry this was not clear. We included the prior specification [$\beta \sim \text{Normal}(0,5)$] in Appendix X on line 59. We have now included this in the main text on line xx.}

3. My biggest concern is that the statistical tests don't seem to test the HRF because they treat year/time as a "nuisance" variable that adds noise to the observations. But the HRF is explicitly about time and how the response of ecosystem function changes over time. So incoporating time as an explicit part of the analysis would be necessary. The easiest way would be to regress ANPP against VWC and allow the slope to vary by year to see if the slopes follow the pattern dictated by HRF (or could use a hierarchical gaussian process/LOWESS model or other non-parametric model to allow for non-lienarity along the VWC gradient - drought - control - irrigation). Perhaps another approach would be to calculate a 'sensitivity' for each replicate (or treatment) within a Bayesian model, then look at how sensitivity to drought/irrigation changes through time (in the same model, i.e. plot-level sensitivity nested within year, with fixed effects to contrast sensitivity to drought/irrigation, a la Wilcox et al. 2017). That is an explicit test of HRF, where the sensitivites could show any number of temporal patterns outlined by Smith.

\response{The reviewer is correct that our analysis is not an explicit test of the Hierarchical Response Framework because we did not explicitly model the effect of time on ecosystem functional response. Unfortunately, we are unable to do so given our data -- in each year every plot shares the same average soil moisture for its particular treatment, making it difficult to fit year-specific slopes for each treatment level.}

\response{That said, the reviewer made an excellent suggestion of conducting a sensitivity analysis \emph{a la} Wilcox et al. 2017, which we have included (see lines xxx - xxx in the Methods section; new results on lines xxx - xxx; and new discussion items on lines xxx-xxx).
The sensitivity analysis allows us to examine how the effect of treatment changes over time, adding another dimension to our analysis that better connects our findings to the Hierachical Response Framework.
We found some evidence that the ecosystem's sensitivity to drought increased over time.
This increase was not large enough to significantly impact ecosystem functional response, and our community composition analysis does not indicate changes in the final years of our experiment.
However, as we note in the final line of our paper (lines xxx-xxx), the increased sensitivity in the latter years of the experiment may portend a coming shift in ecosystem functional response.}

### 2017-12-10 manuscript/review_response_R1.Rmd

+ Several of the dominant species (e.g., threetip sagebrush and bluebunch wheatgrass) have a lifespan that is much longer than the 5-year experiment; while other species in the community have shorter reproductive cycles. It seems that even with a 5-year experiment the longer-lived species of the experimental community may not have had sufficient time to (fully) respond and make it possible to observe measurable changes in the community composition/response.

\response{Yes, we agree that the rate of community turnover is likely related to the lifespans of constituent species. We have added text discussing the impact of lifespan on community turnover in our Conclusion (lines xxx-xxx).}

+ I would find it interesting if the plant community analysis is additionally carried out separately for short-lived vs long-lived species (Fig. 4) -- based on the expectations that short-lived species have more `opportunities' to respond/turnover than long-lived species.

\response{This is a great suggestion, and one we had not thought of. We conducted the analysis suggested by the reviewer and found that community composition was also stable when we restricted the analysis to annual plants (short-lived). In a few years we detected signficant treatment effects on dispersion or mean of community composition. But upon further investigation, we found that these statistical results were detecting changes in rare species because the dominant annuals showed similar dynamics through time, regardless of treatment (Fig. Ax-x). Likewise, the stable perennial community contributes most to ANPP is this systems. We briefly mention this analysis in our Results section on lines xxx-xxx, and show the full analysis in Appendix X.}

@@

Lines 289-299: strong discussion and support for the sizes of the treatment applied. Good work.

\response{Thanks for the kind words.}

Lines 320-325: seems like authors are indicating that ecosystem response was driven by the response of the dominant species. Might be good to bring in some other work that has found similar results and discuss it a bit, for example: ``Drivers of Variation in Aboveground Net Primary Productivity and Plant Community Composition Differ Across a Broad Precipitation Gradient'' (La Pierre et al., 2016).

\response{Thanks for bringing this paper to our attention. We have included a new sentence that references this citation on line xxx in the Discussion.}

### 2017-12-10 manuscript/Appendix1.Rmd
more tiny revisions and matching up with response letter

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \linespread{1.25}
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
  - \usepackage{mathtools}
bibliography: ~/Dropbox/Bibliography/CoexistStability.bib
csl: components/ecology_letters.csl

@@

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A1-\arabic{figure}}  
\renewcommand\thetable{A1-\arabic{table}}  
\renewcommand\thesection{Section A1.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 1}} \\
A.T. Tredennick, A.R. Kleinhesselink, B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe''  \\
\emph{PeerJ}
\end{center}

\linenumbers

\section{Estimating ANPP}
We used a radiometer to nondestructively estimate aboveground net primary productivity.
Our approach relies on relating greenness in a plot to aboveground biomass.
In each year we recorded ground reflectances at four bands, two associated with the red spectrum and two associated with the near-infrared spectrum (Table A1-1).
We took four readings per plot that were averaged for each band.
Bands 1 and 3 correspond to wavelengths collected by the MODIS satellite and bands 2 and 4 correspond to wavelengths collected by the AVHRR satellite.

| Band number | Spectrum name | Center wavelength | Corresponding satellite |
| ----------- | ------------- | ----------------- | ----------------------- |
| 1 | red | 626 nm | AVHRR |
| 2 | red | 652 nm | MODIS |
| 3 | near-infrared | 875 nm | AVHRR |
| 4 | near-infrared | 859 nm | MODIS |

Table: Radiometer specifications.

Using the RED and NIR reflectance values, we calculate the normalized difference vegetation index (NDVI) for each plot based on both AVHRR- and MODIS-based wavelengths.
We calculated NDVI as:
\vspace{-2em}
\begin{align}
\text{NDVI}_{\text{AVHRR}} &= \frac{b_3 - b_1}{b_3 + b_1} \\[10pt]
\text{NDVI}_{\text{MODIS}} &= \frac{b_4 - b_2}{b_4 + b_2}
\end{align}

\noindent{} where $b_x$ refers to band *x* (*x* = 1,2,3,4) in Table A1-1.

To convert plot NDVI to biomass, we regressed known biomass values from calibration plots against NDVI calculate for those plots.
Calibration plots were located near our experiment plots, and each year we located a new set of 10 plots in which we clipped all aboveground biomass, dried it to a constant weight at 60$^\circ$ C, and the weighed.
We used these biomass values to estimate regression parameters for both AVHRR- and MODIS-based NDVI.
We assessed model fit using $R^2$ and, for each year, we used the regression parameters associated with the best fit model to estimate biomass in the experimental plots based on their NDVI values (Table A1-2).
R code for this procedure is in the file "calibrate_radiometer_by_year.R" in the supplemental code set.

```{r ndvi_table, include=T, echo=F, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
library(xtable)
library(tidyverse)
library(dplyr)
tab_dat <- read.csv("../data/estimated_biomass/ndvi_biomass_regression_parameters.csv", row.names = 1) %>%
  filter(year > 2011 & year < 2017)
colnames(tab_dat) <- c("Year", "Intercept", "NDVI Slope", "$R^2$", "Min(biomass)", "Max(biomass)", "\\emph{N}", "Algorithm")

my_caption <- "Details of regression models used to estimate biomass each year."
print(xtable(tab_dat, caption = my_caption, digits = 2),
      caption.placement="top",
      include.rownames = F,
      sanitize.colnames.function = identity,
      comment = FALSE,
      size = "normalsize")
```

>>>>

---
layout: 11pt
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \usepackage{todonotes}
  - \onehalfspacing
  - \usepackage{rotating}
  - \usepackage{color, soul}
  - \usepackage[font={small},labelfont={bf},labelsep=quad]{caption}
  - \usepackage{tikz}
  - \usepackage{bm,mathrsfs}
  - \usepackage[sc]{mathpazo}
  - \usepackage{float}
bibliography: ~/Dropbox/Bibliography/usses_water.bib
csl: components/ecology.csl

@@

\newcommand{\tikzcircle}[2][red,fill=red]{\tikz[baseline=-0.5ex]\draw[#1,radius=#2] (0,0) circle ;}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\renewcommand\thefigure{A1-\arabic{figure}}
\renewcommand\theequation{A1.\arabic{equation}}
\renewcommand\thetable{A1-\arabic{table}}  
\renewcommand\thesection{Section A1.\arabic{section}}

\begin{center}
\textbf{\Large{Appendix 1}} \\
A.T. Tredennick, A.R. Kleinhesselink, J.B. Taylor \& P.B. Adler \\
``Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe'' \\
\emph{PeerJ}
\end{center}

# Details on the plant community and dynamics

Here we provide more details on the plant community in terms of dominance and rarity.
Averaging across time, *Artemisia tripartita* and *Balsamorhiza sagittata* are the two most dominant species in each treatment.
Combined, these two species represent 28\% of total cover in control plots, 25\% of total cover in drought plots, and 25\% of total cover in irrigation plots.
Four to five species dominate the community in general (Figure A1-1), indicating a high level of dominance in this plant community.

We also conducted our community composition analysis with only annual species.
Annual species are shorter-lived than the perennial species in our community, so they may respond more quickly to alterations of precipitation.
In general, our results for annual species conform to the results from the full community analysis in the main text.
Annual plant community composition is relatively stable through time (Fig. A1-6) and in most years there is no evidence that treatment differentiates community composition (Table A1-3).
Note that in some years the ```vegan::metaMDS()``` returned unreliable estimates of Bray-Curtis distances for the annual community because of lack of sufficient data (i.e., many annual species with 0 abundance).

\newpage{}
# Tables

```{r get_data, echo=FALSE, message=FALSE, warning=FALSE}
####
####  LOAD PACKAGES ----
####
library(tidyverse) # all the good things from Wikham
library(dplyr)     # data munging
library(stringr)   # working with strings
library(ggthemes)  # pleasing ggplot2 themes
library(vegan)     # NMDS and community analysis
library(xtable)    # Generating LaTeX tables in RMarkdown

####
####  SET DIRECTORIES; READ IN DATA ----
####
# Set paths
biomass_path <- "../data/estimated_biomass/"
weather_path <- "../data/weather/"
output_path  <- "../output/"
figure_path  <- "../figures/"

# Read in data
plot_biomass <- readRDS(paste0(biomass_path,"permanent_plots_estimated_biomass.RDS"))
weather_data <- read.csv(paste0(weather_path,"ClimateIPM.csv"))
plot_info    <- read.csv("../data/estimated_biomass/quad_info.csv")

suppressWarnings( # ignore factor to character warning. It's OK.
  cover_species <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
    group_by(quad, year, species) %>%
    summarise(total_area = sum(area)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    group_by(year, species, Treatment) %>%
    summarise(avg_area = mean(total_area)) %>%
    ungroup()
)

####
####  COMPARE COMMUNITY COMPOSITION AMONG TREATMENTS ----
####  USING STANDARDIZED ABUNDANCES FROM COVER AND DENSITY
####
suppressWarnings( # ignore factor to character warning. It's OK.
  cover_community_matrix <- read.csv("../data/vital_rates/allrecords_cover_v24.csv") %>%
    group_by(quad, year, species) %>%
    summarise(total_area = sum(area)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    spread(species,total_area, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group) %>%
    ungroup()
)

suppressWarnings( # ignore factor to character warning. It's OK.
  density_community_matrix <- read.csv("../data/vital_rates/allrecords_density_v24.csv") %>%
    mutate(ind_num = 1) %>%
    group_by(year, quad, species) %>%
    summarise(total_inds = sum(ind_num)) %>%
    left_join(plot_info, by = "quad") %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    spread(species,total_inds, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group) %>%
    ungroup()
)

density_spp <- colnames(density_community_matrix[4:ncol(density_community_matrix)])
cover_spp   <- colnames(cover_community_matrix[4:ncol(cover_community_matrix)])
overlap_spp <- density_spp[which(density_spp %in% cover_spp)]
rmids       <- which(colnames(density_community_matrix) %in% overlap_spp)
density_community_matrix <- density_community_matrix[,-rmids]

suppressWarnings( # ignore factor to character warning. It's OK.
  annual_community_matrix <- read.csv("../data/vital_rates/idaho_annuals.csv") %>%
    select(-Notes) %>%
    gather(species,abundance,-c(1:3)) %>%
    left_join(plot_info, by="quad") %>%
    select(-QuadName.y) %>%
    rename(QuadName = QuadName.x) %>%
    filter(Treatment == "Control" | Treatment == "Irrigation" | Treatment == "Drought") %>%
    filter(!str_detect(QuadName, 'P1|P7')) %>%
    filter(year > 2010 & is.na(species)==F) %>%
    separate(species,c("genus","species1")) %>%
    unite(species,genus,species1,sep = " ") %>%
    spread(species,abundance, fill = 0) %>%
    select(-QuadName,-Grazing,-paddock,-Group,-`Lappula occidentalis`,-`Tragopogon dubius`) %>%
    ungroup()
)

annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  for(doyr in unique(annual_community_matrix$year)){
    tmp <- annual_community_matrix %>%
      gather(species, density, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_density = mean(density)) %>%
      arrange(-mean_density) %>%
      mutate(mean_rank = seq_along(mean_density))

    annual_data <- rbind(annual_data, tmp)
  }
}

dom_annuals <- annual_data %>%
  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
  select(-mean_rank) %>%
  left_join(weather_data, by = "year")
```

Table A1-1: Statistical results from linear model relating density of Allysum desertorum to precipitation.
```{r report_models, echo = FALSE}
summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Allyssum desertorum")))
```

\newpage{}
Table A1-2: Statistical results from linear model relating density of Bromus tectorum to precipitation.
```{r report_models2, echo = FALSE}
summary(lm(mean_density~ppt1*Treatment, data = filter(dom_annuals, species == "Bromus tectorum")))
```

\newpage{}

```{r annual_nmds, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
####
####  ANALYZE JUST THE ANNUALS ----
####
# Set missing species abundances as 0, which == absent
annual_community_matrix[is.na(annual_community_matrix)] <- 0

annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))

spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
  select(-ids_to_remove)

nmds_df_annuals <- {}
out_stats <- {}
for(doyr in unique(annual_community_matrix_scaled$year)){
  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
   invisible( capture.output( # already checked for convergence
    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
  ))
  nmds_points <- as.data.frame(tmp_bray$points)
  nmds_points$Treatment <- tmp$Treatment
  nmds_points$Year <- doyr
  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
                   as.numeric(betadisper_stats))
  row.names(tmp_out) <- c()
  out_stats <- rbind(out_stats, data.frame(year = doyr,
                                           test = c("adonis","betadisper"),
                                           n = nrow(tmp),
                                           tmp_out))
}
```

\setcounter{table}{2}

```{r annual_nmds_table, echo = FALSE, include=TRUE, comment=FALSE,warning=FALSE,message=FALSE,results='asis'}
colnames(out_stats) <- c("Year", "Test", "n", "d.f.","Sums","Means","\\emph{F}", "\\emph{P}")
out_stats$d.f. <- as.integer(out_stats$d.f.)
out_stats <- arrange(out_stats, Year)

# Make the table
inyear_caption <- "Results from statistical tests for clustering and dispersion of community composition among precipitation treatments for annual species only. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \\emph{P} values greater than 0.05 indicate there is no support that the treatments differ."
print(xtable(dplyr::select(out_stats, -Sums, -Means), caption = inyear_caption, digits = 2),
       caption.placement="top",
       include.rownames = F,
       sanitize.colnames.function = identity,
       comment = FALSE,
       size = "normalsize")
```

\newpage{}

# Figures

```{r rac_plots, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for perennial species. Basal area of individuals was summed withnin years and plots, and then the total area values were averaged across years and plots for each treatment.", fig.pos="h",fig.align="center"}
####
####  EXTRACT COVER OF MOST DOMINANT SPECIES ----
####
cover_data <- {}
for(dotrt in unique(cover_community_matrix$Treatment)){
  tmp <- cover_community_matrix %>%
    gather(species, area, -quad, -year, -Treatment) %>%
    filter(Treatment == dotrt) %>%
    group_by(Treatment, species) %>%
    summarise(mean_area = mean(area)) %>%
    arrange(-mean_area) %>%
    mutate(mean_rank = seq_along(mean_area))

  cover_data <- rbind(cover_data, tmp)
}

ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(cover_data, mean_rank < 11), aes(x = mean_rank, y = mean_area, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,0.25))+
  xlab("Species rank")+
  ylab(expression(paste("Mean basal area (",m^2,")")))+
  facet_wrap(~Treatment, ncol = 3)

cover_doms <- cover_data %>%
  filter(mean_rank < 3) %>%
  group_by(Treatment) %>%
  summarise(total_area = sum(mean_area))

```

```{r annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Rank abundance curves for annual species. Density of individuals is averaged across years and plots for each treatment."}
annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  tmp <- annual_community_matrix %>%
    gather(species, density, -quad, -year, -Treatment) %>%
    filter(Treatment == dotrt) %>%
    group_by(Treatment, species) %>%
    summarise(mean_density = mean(density)) %>%
    arrange(-mean_density) %>%
    mutate(mean_rank = seq_along(mean_density))

  annual_data <- rbind(annual_data, tmp)
}

ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,200))+
  xlab("Species rank")+
  ylab("Mean density")+
  facet_wrap(~Treatment, ncol = 3)
```

```{r time_perr_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for perennial species in each treatment. Values of mean area were averaged over plots. The four most dominant species are labelled in each panel."}
cover_data <- {}
for(dotrt in unique(cover_community_matrix$Treatment)){
  for(doyr in unique(cover_community_matrix$year)){
    tmp <- cover_community_matrix %>%
      gather(species, area, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_area = mean(area)) %>%
      arrange(-mean_area) %>%
      mutate(mean_rank = seq_along(mean_area))

  cover_data <- rbind(cover_data, tmp)
  }
}
ggplot(cover_data, aes(x = mean_rank, y = mean_area))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(cover_data, mean_rank < 5), aes(x = mean_rank, y = mean_area, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,0.25))+
  xlab("Species rank")+
  ylab(expression(paste("Mean basal area (",m^2,")")))+
  facet_grid(Treatment~year)
```

```{r time_annual_racs, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=6, fig.cap="Time series of rank abundance curves for annual species in each treatment. Values of density were averaged over plots. The two most dominant species are labelled in each panel."}
annual_data <- {}
for(dotrt in unique(annual_community_matrix$Treatment)){
  for(doyr in unique(annual_community_matrix$year)){
    tmp <- annual_community_matrix %>%
      gather(species, density, -quad, -year, -Treatment) %>%
      filter(Treatment == dotrt, year == doyr) %>%
      group_by(Treatment, year, species) %>%
      summarise(mean_density = mean(density)) %>%
      arrange(-mean_density) %>%
      mutate(mean_rank = seq_along(mean_density))

    annual_data <- rbind(annual_data, tmp)
  }
}

ggplot(annual_data, aes(x = mean_rank, y = mean_density))+
  geom_line()+
  geom_point(size = 2)+
  geom_text(data=filter(annual_data, mean_rank < 3), aes(x = mean_rank, y = mean_density, label = species), 
            hjust = -0.05, nudge_x = 0.2, angle = 45, size = 2)+
  scale_y_continuous(limits = c(0,400))+
  scale_x_continuous(breaks = c(1,6,11))+
  xlab("Species rank")+
  ylab("Mean density")+
  facet_grid(Treatment~year)
```

\newpage{}

```{r annual_weather, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Relationship between precipitation and abundance of dominant annuals."}
dom_annuals <- annual_data %>%
  filter(species %in% c("Allyssum desertorum","Bromus tectorum")) %>%
  select(-mean_rank) %>%
  rename(Species = species) %>%
  left_join(weather_data, by = "year")

ggplot(dom_annuals, aes(x = ppt1, y = mean_density, color = Species))+
  geom_point()+
  xlab("Precipitation (mm)")+
  ylab("Abundance")
```

```{r annual_nmds2, echo=FALSE, include=TRUE, message=FALSE, warning=FALSE, fig.height=3, fig.cap="Nonmetric multidimensional scaling scores representing annual plant communities in each plot, colored by treatment."}
####
####  ANALYZE JUST THE ANNUALS ----
####
# Set missing species abundances as 0, which == absent
annual_community_matrix[is.na(annual_community_matrix)] <- 0

annual_community_matrix_scaled <- cbind(annual_community_matrix[,1:3,drop=F],
                                        scale(annual_community_matrix[,4:ncol(annual_community_matrix)]))

spp_to_remove <- names(which(is.nan(colMeans(annual_community_matrix_scaled[4:ncol(annual_community_matrix_scaled)]))))
ids_to_remove <- which(colnames(annual_community_matrix_scaled) %in% spp_to_remove)
annual_community_matrix_scaled <- annual_community_matrix_scaled %>%
  select(-ids_to_remove)

nmds_df_annuals <- {}
out_stats <- {}
for(doyr in unique(annual_community_matrix_scaled$year)){
  tmp <- filter(annual_community_matrix_scaled, year == doyr) %>% select(-year)
   invisible( capture.output( # already checked for convergence
    tmp_bray <- metaMDS(tmp[3:ncol(tmp)]+2, "bray", trymax = 999)
  ))
  nmds_points <- as.data.frame(tmp_bray$points)
  nmds_points$Treatment <- tmp$Treatment
  nmds_points$Year <- doyr
  nmds_df_annuals <- rbind(nmds_df_annuals, nmds_points)
  adonis_stats <- as.data.frame(vegan::adonis(tmp[3:ncol(tmp)]+2~tmp$Treatment)$aov.tab)[1,]
  betadisper_stats <- anova(vegan::betadisper(vegan::vegdist(tmp[3:ncol(tmp)]+2),tmp$Treatment))[1,]
  tmp_out <- rbind(adonis_stats[,-which(colnames(adonis_stats)=="R2")],
                   as.numeric(betadisper_stats))
  row.names(tmp_out) <- c()
  out_stats <- rbind(out_stats, data.frame(year = doyr,
                                           test = c("adonis","betadisper"),
                                           n = nrow(tmp),
                                           tmp_out))
}

mycols <- c("#009E73", "#0072B2", "#D55E00")
ggplot(nmds_df_annuals, aes(x=MDS1, y=MDS2, color=Treatment))+
  geom_point()+
  geom_point(color="grey35",shape=1)+
  scale_color_manual(values=mycols)+
  ylab("NMDS 2")+
  xlab("NMDS 1")+
  facet_wrap(~Year)+
  #scale_y_continuous(limits=c(-0.15,0.15))+
  #scale_x_continuous(limits=c(-0.18,0.18))+
  coord_fixed()
```
